---
title: "Extracting the Estimates from the five models"
author: "Oluwayomi Akinfenwa"
date: "30/07/2024"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
  bookdown::html_document2:
    toc: no
    toc_float: yes
    toc_depth: 4
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.duplicate.label = 'allow')
```

Sourcing the file containing all the data and libraries needed for this analysis from the Rmd file named: 01_PISA_Analysis_Data.Rmd.
This file contains the raw data from OECD website and all the necessary data for this analysis.

```{r loading the data}
sys.source("00_PISA_Data.R", envir = knitr::knit_global())
```

## Loading the hierarchical models from the saved R.Data file

```{r}
#Independent-Country specific model
load(here::here("Manuscript_Models", "CountryInd_BRMSModel.Rdata"))

#Country specific model
load(here::here("Manuscript_Models", "Country_BRMSModel.Rdata"))

#Region hierarchical model
load(here::here("Manuscript_Models", "Region_BRMSModel.Rdata"))

#Income hierarchical model
load(here::here("Manuscript_Models", "Income_BRMSModel.Rdata"))

#Income-Region hierarchical model
load(here::here("Manuscript_Models", "Income_region_BRMSModel.Rdata"))
```

### Pulling the estimates from all the fitted models

#### Estimates from the independent-country-specific model

```{r}
#The global estimates is the estimate for Albania. Albania is used as the base country.
# also unitedkingdom was merged together.
#hence, the str_replace_all might not work to fix this and I have to fix this manually.

CountryInd_draws <-posterior::as_draws_rvars(CountryInd_BRMSModel, 
                variable = "^b_Country|^b_Intercept",regex = TRUE) |>
  purrr::map_dfr(as_tibble, .id = "name")|>
  mutate(rvar=value+value[1]) # Values are the offsets from the global estimates, hence, we add the offset to the global estimates.

#Albania is used as the base country, hence the intercept estimate (b_Intercept) is the rvar for Albania.
CountryInd_draws$rvar[1] <- CountryInd_draws$value[1]
  
CountryInd_I <- CountryInd_draws |>
  mutate(Country = stringr::str_replace_all(name, 
                          stringr::fixed("b_Country"), ""))|>
  mutate(Country = stringr::str_replace_all(Country, 
                          stringr::fixed("b_Intercept"), "Albania"))|>
  mutate(Country = stringr::str_replace(Country, 
                          stringr::fixed("UnitedKingdom"), "United Kingdom"))|>
  right_join(SNames, join_by(Country== Country))|> #joining the country and the hierarchical groupings with the model estimates
    mutate(term = "Intercept") |>
  select(Country, term, rvar)

#Slope estimates
CountrySlop_draws <- posterior::as_draws_rvars(CountryInd_BRMSModel, 
                    variable = "^b_year_orig", regex = TRUE) |>
  purrr::map_dfr(as_tibble, .id = "name")|>
  mutate(rvar=value+value[1])

#Albania is used as the base country, hence the intercept estimate is the rvar for Albania.
CountrySlop_draws$rvar[1] <- CountrySlop_draws$value[1]

CountryInd_S <- CountrySlop_draws |>
  mutate(Country = stringr::str_replace_all(name, stringr::fixed("b_year_orig:Country"), "")) |>
  mutate(Country = stringr::str_replace(Country, stringr::fixed("b_year_orig"), "Albania")) |>
  mutate(Country = stringr::str_replace(Country, stringr::fixed("UnitedKingdom"), "United Kingdom")) |>
  right_join(SNames, join_by(Country== Country))|>
  mutate(term = "year_orig") |>
  select(Country, term, rvar)

CountryInd_IS <- rbind(CountryInd_I, CountryInd_S)

## Tidying the results
CountryInd_IS_Est <- CountryInd_IS |> mutate(coef=median(rvar))|> 
  select(-rvar)|>
  pivot_wider(values_from=coef, names_from=term)

#To identify the estimates, we assign a new variable called model with the model name across each model estimate.
CountryInd_Int_Est <- CountryInd_I |>
  mutate(Model = "Independent \nModel") |>
  select(Country, Model, rvar)

CountryInd_Slop_Est <- CountryInd_S|>
  mutate(Model = "Independent \nModel") |>
  select(Country, Model, rvar)

# CountryInd_Int_Est is the Intercept estimates from the independent model while
# CountryInd_Slop_Est is the Slope estimates from the independent model.
```

#### Estimates from the country-specific model

```{r}
## Pulling the intercept results. The global intercept (b_Intercept) and the country offsets.
Country_I <- spread_rvars(Country_BRMSModel, r_Country[Country,term], b_Intercept) |>
  filter(term == "Intercept") |>
  mutate(Country = stringr::str_replace_all(Country, fixed("."), " ")) |>
  mutate(r_Country = r_Country+ b_Intercept) |>
  rename(rvar=r_Country)

Country_Int <- Country_I |>
  select(Country, term, rvar)

## Pulling the slope results. The global slope (b_year_orig) and the country offsets.
Country_S <- spread_rvars(Country_BRMSModel, r_Country[Country,term], b_year_orig) |>
  filter(term == "year_orig") |>
  mutate(Country = stringr::str_replace_all(Country, fixed("."), " ")) |>
  mutate(r_Country = r_Country+ b_year_orig) |>
   rename(rvar=r_Country)

Country_Slop <- Country_S|>
  select(Country, term, rvar)

## Tidying the estimates
Country_IS <- rbind(Country_Int, Country_Slop)

Country_IS_Est <- Country_IS |> mutate(coef=median(rvar))|> 
  select(-rvar)|>
  pivot_wider(values_from=coef, names_from=term)

#The hierarchical estimate for the country-specific model is the b_intercept (global estimate)
Country_Int_est <- Country_I |>
  mutate(Model = "Country \nModel") |>
  select(c(Country, Model, rvar = b_Intercept))

Country_Slop_est <- Country_S |>
  mutate(Model = "Country \nModel") |>
  select(c(Country, Model, rvar= b_year_orig))

#Country-level estimates from the country-specific model
## Intercept Estimates
Country_Int_Est <- Country_I |>
  mutate(Model = "Country \nModel") |>
  select(c(Country, Model, rvar))

## Slope Estimates
Country_Slop_Est <- Country_S |>
  mutate(Model = "Country \nModel") |>
  select(c(Country, Model, rvar))
```

#### Estimates from the Region hierarchical model

```{r}
Region_int <- spread_rvars(Region_BRMSModel, r_Region[group1,term], b_Intercept) |>
  filter(term == "Intercept") |>
  mutate(group1 = stringr::str_replace_all(group1, fixed("."), " ")) |>
  mutate(r_Region = r_Region+ b_Intercept) |>
  rename(rvar=r_Region, Region = group1) |>
  select(Region, term, rvar)

Region_slop <- spread_rvars(Region_BRMSModel, r_Region[group1,term], b_year_orig) |>
  filter(term == "year_orig") |>
  mutate(group1 = stringr::str_replace_all(group1, fixed("."), " ")) |>
  mutate(r_Region = r_Region+  b_year_orig) |>
  rename(rvar=r_Region, Region = group1)|>
  select(Region, term, rvar)

#Extending the respective regional estimates to the countries
Region_Int <- Region_int |> 
  right_join(select(Names, Country, Region),  
             join_by(Region == Region))

Region_Slop <- Region_slop |> 
  right_join(select(Names, Country, Region),  
             join_by(Region == Region))
  
## The country-level estimates from the region model
Rcountry_Int <- spread_rvars(Region_BRMSModel, r_Region[group1,term], r_Country[group,term], b_Intercept) |> # group1 is region and group is country
  filter(term == "Intercept") |>
  mutate(group1 = stringr::str_replace_all(group1, fixed("."), " "), 
         group = stringr::str_replace_all(group, fixed("."), " ")) |>
  right_join(select(Names, Country, Region),
             join_by(group==Country,group1==Region)) |> #this join everything from the Names data set with its intercession from the model result
  mutate(r_Country =  r_Country+ r_Region+ b_Intercept) |> # Adding the global estimate with the region offset and the country offset
  rename(rvar=r_Country, Region=group1, Country =group) |>
  select(Country, Region, term, rvar)

Rcountry_Slop <- spread_rvars(Region_BRMSModel, r_Region[group1,term],r_Country[group,term],b_year_orig) |>
  filter(term == "year_orig") |>
  mutate(group1 = stringr::str_replace_all(group1, fixed("."), " "), 
         group = stringr::str_replace_all(group, fixed("."), " ")) |>
  right_join(select(Names, Country, Region),
             join_by(group==Country,group1==Region))|>
  mutate(r_Country =  r_Country+ r_Region+ b_year_orig) |>
  rename(rvar=r_Country, Region=group1, Country =group) |>
  select(Country, Region, term, rvar)

Rcountry_IS <- rbind(Rcountry_Int, Rcountry_Slop) 
```

### Tidying the estimates from the Region hierarchical model

```{r}
#median estimates
Rcountry_IS_Est <- Rcountry_IS |> mutate(coef=median(rvar))|> 
  select(-rvar)|>
  pivot_wider(values_from=coef, names_from=term)

## Credible interval
#Hierarchy Estimates
## Intercept estimates
Region_Int_est <- Region_Int |>
  mutate(Model = "Region \nModel")|>
  select(Country, Model, rvar)
## Slope estimates
Region_Slop_est <- Region_Slop |>
  mutate(Model = "Region \nModel")|>
  select(c(Country, Model, rvar))

#Country Estimates
## Intercept estimates
Region_Int_Est <- Rcountry_Int |> 
  mutate(Model = "Region \nModel")|>
  select(c(Country, Model, rvar))
## Slope Estimates
Region_Slop_Est <- Rcountry_Slop |> 
  mutate(Model = "Region \nModel")|>
  select(Country, Model, rvar)
```

#### Estimates from the income hierarchical model

```{r}
Income_int <- spread_rvars(Income_BRMSModel, r_Income[group1,term],b_Intercept) |>
  filter(term == "Intercept") |>
  mutate(group1 = stringr::str_replace_all(group1, fixed("."), " ")) |>
  mutate(r_Income = r_Income+ b_Intercept) |>
  rename(rvar=r_Income, Income= group1) |>
  select(Income, term, rvar)

Income_slop <- spread_rvars(Income_BRMSModel, r_Income[group1,term], b_year_orig) |>
  filter(term == "year_orig") |>
  mutate(group1 = stringr::str_replace_all(group1, fixed("."), " ")) |>
  mutate(r_Income = r_Income+  b_year_orig) |>
  rename(rvar=r_Income, Income= group1)|>
  select(Income, term, rvar)

#Extending the income hierarchical estimates across the countries.
Income_Int <- Income_int |>
  right_join(select(Names, Country, Income),
             join_by(Income==Income))

Income_Slop <- Income_slop |>
  right_join(select(Names, Country, Income),
             join_by(Income==Income))
```

pulling the intercepts and slopes for each country based on the influence of the Income hierarchy

```{r}
Icountry_Int <- spread_rvars(Income_BRMSModel, r_Income[group1,term],r_Country[group,term],b_Intercept) |>
  filter(term == "Intercept") |>
  mutate(group1 = stringr::str_replace_all(group1, fixed("."), " "), 
         group = stringr::str_replace_all(group, fixed("."), " ")) |>
  right_join(select(Names, Country, Income),
             join_by(group==Country, group1==Income))|> #this joins the Unique country names and income with their intercession from the model result
  mutate(r_Country =  r_Country+ r_Income+ b_Intercept) |>
  rename(rvar=r_Country, Country = group, Income = group1) |>
  select(Country, Income, term, rvar)

Icountry_Slop <- spread_rvars(Income_BRMSModel, r_Income[group1,term],r_Country[group,term],b_year_orig) |>
  filter(term == "year_orig") |>
  mutate(group1 = stringr::str_replace_all(group1, fixed("."), " "), 
         group = stringr::str_replace_all(group, fixed("."), " ")) |>
  right_join(select(Names, Country, Income),
             join_by(group==Country,group1== Income))|>
  mutate(r_Country =  r_Country+ r_Income+ b_year_orig) |>
  rename(rvar=r_Country, Country = group, Income = group1) |>
  select(Country, Income, term, rvar)

Icountry_IS <- rbind(Icountry_Int, Icountry_Slop) 
```

#### Tidying the estimates from the income model

```{r, fig.width=10, fig.height=8}
#Country-estimates based on the hierarchical pulling.
Icountry_IS_Est <- Icountry_IS |> mutate(coef=median(rvar))|> 
  select(-rvar)|>
  pivot_wider(values_from=coef, names_from=term)

## Credible interval
#Hierarchy Estimates
## Intercept estimates
Income_Int_est <- Income_Int |> 
  mutate(Model = "Income \nModel")|>
  select(c(Country, Model, rvar))
## Slope estimates
Income_Slop_est <- Income_Slop |> 
  mutate(Model = "Income \nModel")|>
  select(c(Country, Model,rvar))

#Country Estimates
## Intercept estimates
Income_Int_Est <- Icountry_Int |> 
  mutate(Model = "Income \nModel")|>
  select(Country, Model, rvar)
## Slope estimates
Income_Slop_Est <- Icountry_Slop |> 
  mutate(Model = "Income \nModel")|>
  select(Country, Model,rvar)
```

#### Estimates from the Income_region hierarchical model

```{r}
Income_region_int <- spread_rvars(Income_region_BRMSModel, r_Income_region[group1,term],b_Intercept) |>
  filter(term == "Intercept") |>
  mutate(group1 = stringr::str_replace_all(group1, fixed("."), " ")) |>
  mutate(r_Income_region = r_Income_region+ b_Intercept) |>
  rename(rvar=r_Income_region, `Income_region` = group1) |>
  select(Income_region, term, rvar)

Income_region_slop <- spread_rvars(Income_region_BRMSModel, r_Income_region[group1,term], b_year_orig) |>
  filter(term == "year_orig") |>
  mutate(group1 = stringr::str_replace_all(group1, fixed("."), " ")) |>
  mutate(r_Income_region = r_Income_region+ b_year_orig) |>
  rename(rvar=r_Income_region, `Income_region` = group1) |>
  select(Income_region, term, rvar)

#Extending the income-region hierarchical estimates across the countries.
Income_region_Int <- Income_region_int |>
    right_join(select(Names, Country, Income_region),
             join_by(Income_region==Income_region))

Income_region_Slop <- Income_region_slop |>
    right_join(select(Names, Country, Income_region),
             join_by(Income_region==Income_region))
```

pulling the intercepts and slopes for each country based on the influence of the Income-region hierarchy.

```{r}
IRCountry_Int <- spread_rvars(Income_region_BRMSModel, r_Income_region[group1,term],r_Country[group,term],b_Intercept) |>
  filter(term == "Intercept") |>
  mutate(group1 = stringr::str_replace_all(group1, fixed("."), " "), 
         group = stringr::str_replace_all(group, fixed("."), " ")) |>
  right_join(select(Names, Country, Income_region),
             join_by(group==Country, group1==Income_region))|> #this joins the Unique country names and income with their intercession from the model result
  mutate(r_Country =  r_Country+ r_Income_region+ b_Intercept) |>
  rename(rvar=r_Country, Country = group, Income_region = group1) |>
  select(Country, Income_region, term, rvar)


IRCountry_Slop <- spread_rvars(Income_region_BRMSModel, r_Income_region[group1,term],r_Country[group,term],b_year_orig) |>
  filter(term == "year_orig") |>
  mutate(group1 = stringr::str_replace_all(group1, fixed("."), " "), 
         group = stringr::str_replace_all(group, fixed("."), " ")) |>
  right_join(select(Names, Country, Income_region),
             join_by(group==Country, group1==Income_region))|> #this joins the Unique country names and income with their intercession from the model result
  mutate(r_Country =r_Country+ r_Income_region+ b_year_orig) |>
  rename(rvar=r_Country, Country = group, Income_region = group1) |>
  select(Country, Income_region, term, rvar)

IRCountry_IS <- rbind(IRCountry_Int, IRCountry_Slop) 
```

#### Tidying the estimates from the model

```{r}
IRCountry_IS_Est <- IRCountry_IS |> mutate(coef=median(rvar))|> 
  select(-rvar)|>
  pivot_wider(values_from=coef, names_from=term)

## Credible interval
#Hierarchy Estimates
## Intercept estimates
Income_region_Int_est <- Income_region_Int |> 
  mutate(Model = "Income-region \nModel")|>
  select(Country, Model, rvar)
## Slope estimates
Income_region_Slop_est <- Income_region_Slop |> 
  mutate(Model = "Income-region \nModel")|>
  select(Country, Model,rvar)

#Country Estimates
Income_region_Int_Est <- IRCountry_Int |> 
  mutate(Model = "Income-region \nModel")|>
  select(Country, Model,rvar)
## Slope estimates
Income_region_Slop_Est <- IRCountry_Slop |> 
  mutate(Model = "Income-region \nModel")|>
  select(Country, Model,rvar)
```

### Plotting the regression model fits on the data

```{r}
if (!dir.exists(here::here("Saved_Plots/Regression_fits")))
 dir.create(here::here("Saved_Plots/Regression_fits"))
```

#### Independent country model fit on the data

```{r}
pdf("Saved_Plots/Regression_fits/Independent_fit.pdf",
    width = 11,
    height = 8)
ggplot(PISA_Europe_Data, aes(x=year_orig, y=math))+ 
  geom_point()+
  facet_wrap(~Country)+
  geom_abline(data=CountryInd_IS_Est, aes(slope=year_orig, intercept= Intercept), color="black")+
  scale_x_continuous(breaks = seq(-15,0,3),labels = function(x) x+2018) +
  xlab("year")+ ggtitle("Independent country model fit of average maths scores over time")+
  theme_bw()+
  theme(panel.spacing=unit(1,"lines"),
        legend.position = "bottom",
        plot.subtitle = element_text(family = "Consolas"),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

dev.off()
```

#### Country model fit on the data

```{r}
pdf("Saved_Plots/Regression_fits/Country_specific_fit.pdf",
    width = 11,
    height = 8)

ggplot(PISA_Europe_Data, aes(x=year_orig, y=math))+ 
  geom_point()+
  facet_wrap(~Country)+
  geom_abline(data=Country_IS_Est, aes(slope=year_orig, intercept= Intercept), color="magenta3")+
  scale_x_continuous(breaks = seq(-15,0,3),labels = function(x) x+2018) +
  xlab("year")+ ggtitle("Country model fit of average maths scores over time")+
  theme_bw()+
  theme(panel.spacing=unit(1,"lines"),
        legend.position = "bottom",
        plot.subtitle = element_text(family = "Consolas"),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))
dev.off()
```

#### Region model fit on the data

```{r}
pdf("Saved_Plots/Regression_fits/Region_hierarchical_fit.pdf",
    width = 11,
    height = 8)
ggplot(PISA_Europe_Data, aes(x=year_orig, y=math))+ 
  geom_point()+
  ggh4x::facet_nested_wrap(vars(Region, Country), nrow = 5)+
  geom_abline(data= Rcountry_IS_Est, aes(slope=year_orig, intercept= Intercept), color="blue3")+
  scale_x_continuous(breaks = seq(-15,0,3),labels = function(x) x+2018) +
  xlab("year")+ ggtitle("Region model fit of average maths scores over time")+
  theme_bw()+
  theme(panel.spacing=unit(1,"lines"),
        legend.position = "bottom",
        plot.subtitle = element_text(family = "Consolas"),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))
dev.off()
```

#### Income model fit on the data

```{r}
pdf("Saved_Plots/Regression_fits/Income_hierarchical_fit.pdf",
    width = 11,
    height = 8)
ggplot(PISA_Europe_Data, aes(x=year_orig, y=math))+ 
  geom_point()+
  ggh4x::facet_nested_wrap(vars(Income, Country), nrow = 5)+
  geom_abline(data= Icountry_IS_Est, aes(slope=year_orig, intercept= Intercept), color="red3")+
  scale_x_continuous(breaks = seq(-15,0,3),labels = function(x) x+2018) +
  xlab("year")+ ggtitle("Income model fit of average maths scores over time")+
  theme_bw()+
  theme(panel.spacing=unit(1,"lines"),
        legend.position = "bottom",
        plot.subtitle = element_text(family = "Consolas"),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))
dev.off()
```

#### IncomeRegion model fit on the data

```{r}
pdf("Saved_Plots/Regression_fits/Income-region_hierarchical_fit.pdf",
    width = 11,
    height = 8)
ggplot(PISA_Europe_Data, aes(x=year_orig, y=math))+ 
  geom_point()+
  ggh4x::facet_nested_wrap(vars(Income_region, Country), nrow = 5)+
  geom_abline(data= IRCountry_IS_Est, aes(slope=year_orig, intercept= Intercept), color="red3")+
  scale_x_continuous(breaks = seq(-15,0,3),labels = function(x) x+2018) +
  xlab("year")+ ggtitle("Income model fit of average maths scores over time")+
  theme_bw()+
  theme(panel.spacing=unit(1,"lines"),
        legend.position = "bottom",
        plot.subtitle = element_text(family = "Consolas"),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))
dev.off()
```


We are interested in sourcing this file rather than re-running the code to create our customised visualisation of multiple models.

```{r}
knitr::purl("01_Models_Estimates.Rmd", output = "01_Models_Estimates.R")
```

