---
title: "Background Section of the Paper"
author: "Oluwayomi Akinfenwa"
date: "06/08/2024"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
  bookdown::html_document2:
    toc: no
    toc_float: yes
    toc_depth: 4
    number_sections: yes
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.duplicate.label = 'allow')
```

Creating a folder to store all the plots

```{r, libraries, include = FALSE}
if (!dir.exists(here::here("Saved_plots")))
  dir.create(here::here("Saved_plots"))
```

Loading all the necessary packages.

```{r}
rm(list = ls()) # clear workspace
library(brms)
library(dplyr)
library(tidybayes)
library(ggplot2)
library(ggragged)
```


**The conventional approach to visualise estimates from a BHM or linear regression model.**

To illustrate the conventional approach, we use the country-specific hierarchical model, where countries borrow strength from a global distribution.

$$ \text{math} \sim \; \text{year} + (1 + \text{year} | \text{Country}) $$

All BRMS models used for this analysis are found in a folder named: 01_Hierarchical_models.Rmd.

The coefficients are saved in 

```{r}

#Country specific model
# These saved files are constructed by running the code in 01_Hierarchical_models.Rmd
model2 <- readRDS(here::here("Manuscript_models", "model2.rds"))
load(here::here("Manuscript_models", "extract_IS.Rdata"))
```


```{r}

countries_IS2 <- filter(countries_IS, modelID=="model2") |>
  rename(Country=group)

Country_IS_Est<-  filter(countries_IS2, modelID=="model2") |> 
  mutate(coef=median(rvar)) |> 
  select(-c(rvar, modelID))|>
  pivot_wider(values_from=coef, names_from=term)

```


#### Visualising this model fit in the data space

```{r}

pdf(here::here("Saved_plots", "Country_specific.pdf"))

baseline_year <- 2018

ggplot(data = model2$data,  # Europe PISA data up to 2018
         aes(x=year_orig, y=math))+
  geom_point(color = "grey25")+
   geom_abline(data= Country_IS_Est, 
               aes(slope=year_orig, intercept= Intercept))+
  facet_wrap(~ Country, nrow = 6,
             labeller = label_wrap_gen(width= 10))+
  labs(x = " ", y = " ") +
 scale_x_continuous(breaks = seq(-15,0,3), labels = function(x) x+baseline_year)+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5,
                                      size = 7, face = "bold"),
        plot.title = element_text(hjust = 0.4, size = 10),
        axis.text.y = element_text(size = 10, face = "bold"),
        strip.text.x = element_text(size = 9, face = "bold"))
dev.off()
```

#### Pulling out the offsets from the country-specific model

```{r}
#The global intercept and slope from model2

global_I <- filter(groups_IS_by_country, modelID=="model2", term=="Intercept")


CI_offset <- countries_IS2 |> 
  filter(term=="Intercept")
CI_offset$rvar <- CI_offset$rvar - global_I$rvar 


# The visual

pdf(here::here("Saved_plots", "offset.pdf"))

ggplot(data = CI_offset)+
  ggdist::stat_pointinterval(aes(xdist = rvar, y= Country),
          color ="burlywood4", .width = c(.50, .95), point_size = 1.4)+
  geom_vline(xintercept = 0, color = "red3") +
  labs(x = " ", y = " ")
dev.off()
```

