---
title: "Model prediction from the Income-region model"
author: "Oluwayomi Akinfenwa"
date: "01/10/2024"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
  bookdown::html_document2:
    toc: no
    toc_float: yes
    toc_depth: 4
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.duplicate.label = 'allow')
```

```{r}
library(dplyr)
library(brms)
library(ggplot2)
library(tidybayes)
```

Sourcing the file containing all the data and country groupings needed for this analysis from 00_PISA_Data.R.


```{r loading the data}
rm(list = ls()) # clear workspace

load(here::here("PISA_data", "Europe_Pisamaths.Rdata")) # Pisa_Europe_Data constructed by sourcing 00_PISA_Data.R
load(here::here("PISA_data", "Country_groups.Rdata")) # Country_groups constructed by sourcing 00_PISA_Data.R

```


## Loading the hierarchical model


```{r}
#Income-region hierarchical model
model5 <- readRDS(here::here("Manuscript_models", "model5.rds")) #saved in the 01_hierarchical_models.Rmd file

```


### Pulling the posterior draws estimates for the year 2022.

We want to calculate the differences between the predicted values and the observed PISA scores for year 2022.

Recall that we didn't include the PISA2022 observations in our modelling, hence we chose the Income_region model to predict PISA2022 scores and visualise the differences between predicted and observed PISA2022.

```{r}

#Creating a grid of the data, make prediction for year 2022, and observed 2022 response
baseline_year <- 2018

Pisa22 <- Pisa_Europe_Data |>
  filter(year==2022) |>
  select(Country, math)

Pred_2022 <- Pisa_Europe_Data |>
  select(-math) |>
  slice(1, .by=Country) |>
  mutate(year=2022, year_orig = year-baseline_year) |>
  add_predicted_rvars(model5) |>
  rename(pred = .prediction) |>
  left_join(Pisa22, by = join_by(Country)) |>
  mutate(math0 = math - pred, prob = posterior::Pr(pred <= math)) |>
  arrange(!is.na(prob), prob) |>
  mutate(Country = factor(Country, levels = Country))
  
```


We want to plot the differences between predicted and observed PISA scores for 2022.


```{r}
#Income-region flag
IR_flag <- Country_groups |>
   slice(1, .by=Income_region) |>
   select(-Country, -Income, -Region) |>
    as.data.frame()
```


#### Visualising the differences between the predicted and observed from the incomeregion model.

```{r}
if (!dir.exists(here::here("Saved_plots")))
  dir.create(here::here("Saved_plots"))

pdf(here::here("Saved_plots", "Pred_Obsv.pdf"))
# The plot
P_IR <- ggplot(data = Pred_2022)+
  geom_vline(xintercept = 0, color = "burlywood4")+
        ggdist::stat_pointinterval(aes(xdist = math0, y= Country),
                  .width = c(.50, .80, .95), point_size = 1.4, color ="grey35")+
  facet_grid(Income_region ~ ., scales = "free_y", space = "free_y", switch = "y")+ 
  labs(x ="Differences between predicted and observed PISA2022 estimates", y = " ")+
  theme_bw()+
  theme(strip.text = element_text(size = 6.3, face = "bold"),
        axis.text.y = element_text(size = 8.0),
        axis.title.x = element_text(size = 8))  # Adjust the x-axis title size


#creating the grid color
G <- ggplotGrob(P_IR) # for other ggplots

stripl <- which(grepl('strip-l', G$layout$name))  #the y axis names has been switch to the left, hence the reason for the strip-l
for (i in stripl){
  j <- which(grepl('rect', G$grobs[[i]]$grobs[[1]]$childrenOrder))
  k <- which(grepl('text', G$grobs[[i]]$grobs[[1]]$childrenOrder))
  grobtext <- G$grobs[[i]]$grobs[[1]]$children[[k]]$children[[1]]$label
  st <- match(grobtext, IR_flag$Income_region)
  if (!is.na(st)){
    G$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- IR_flag[st, "color_region"]
    r <- which(grepl('text', G$grobs[[i]]$grobs[[1]]$children[[k]]$childrenOrder))
    G$grobs[[i]]$grobs[[1]]$children[[k]]$children[[r]]$gp$col <- IR_flag[st, "color_income"]
    G$grobs[[i]]$grobs[[1]]$children[[k]]$children[[1]]$label <- 
      stringr::str_replace_all(grobtext, stringr::fixed("_ "), "\n")
  }
}

grid.draw(G)
dev.off()
```
